---
title: "Automated QA of SPC and SEN data"
author: "Hackathon 2025"
date: "2025-09-04"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      dev = 'svg')


library(tidyverse)
library(janitor)


time_identifier = 202425
options(scipen = 999)
theme_set(theme_minimal())
```

```{r load in the data, include=FALSE}
#read in the data to be used in the report and comparisons
SPC_internal_comparison_summary <- read.csv("./SPC_combined_summary.csv")


SPC_ap_age_national <- read.csv("./national_total_spc_ap_census_age.csv")
SPC_ap_age_regional <- read.csv("./regional_total_spc_ap_census_age.csv")
SPC_ap_age_la <- read.csv("./la_total_spc_ap_census_age.csv")



```

```{r data comparisons and calculations, include=FALSE}
#example code commented out below

# ap_age_national_comparison <- SPC_ap_age_national %>%
#   left_join(SEC_ap_age_national, by = "age") %>%
#   mutate(mismatch = ifelse(national_total.x != national_total.y, 1, 0))

#calcualate percentage all match
SPC_internal_comparison_summary <- SPC_internal_comparison_summary %>%
  mutate(all_match_count = ifelse(Internal_comparison_result == "All match", 1, 0),
         regional_mismatch_count = ifelse(Internal_comparison_result == "regional mismatches", 1, 0),
         national_mismatch_count = ifelse(Internal_comparison_result == "national mismatches", 1, 0),
         regional_national_mismatch_count = ifelse(Internal_comparison_result == "regional & national mismatches", 1, 0))

#count of matching and mismatching
SPC_all_match_total <- sum(SPC_internal_comparison_summary$all_match_count)
SPC_regional_mismatch_count <- sum(SPC_internal_comparison_summary$regional_mismatch_count)
SPC_national_mismatch_count <- sum(SPC_internal_comparison_summary$national_mismatch_count)
SPC_regional_national_mismatch_count <- sum(SPC_internal_comparison_summary$regional_national_mismatch_count)

#percentage matching and mismatching
SPC_all_match_perc <- paste0(as.character(((sum(SPC_internal_comparison_summary$all_match_count) /  nrow(SPC_internal_comparison_summary))*100)),"%")
SPC_regional_mismatch_perc <- paste0(as.character(((sum(SPC_internal_comparison_summary$regional_mismatch_count) /  nrow(SPC_internal_comparison_summary))*100)),"%")
SPC_national_mismatch_prec <- paste0(as.character(((sum(SPC_internal_comparison_summary$national_mismatch_count) /  nrow(SPC_internal_comparison_summary))*100)),"%")
SPC_regional_national_mismatch_prec <- paste0(as.character(((sum(SPC_internal_comparison_summary$regional_national_mismatch_count) /  nrow(SPC_internal_comparison_summary))*100)),"%")

#mismatched rows
SPC_regional_mismatch_rows <- SPC_internal_comparison_summary %>%
  filter(regional_mismatch_count == 1)
SPC_national_mismatch_rows <- SPC_internal_comparison_summary %>%
  filter(national_mismatch_count == 1)
SPC_regional_national_mismatch_rows <- SPC_internal_comparison_summary %>%
  filter(regional_national_mismatch_count == 1)

```


## Internal comparisons

Comparisons have been made internally for assurance that the totals were the same when grouping by Local Authority, Region and Nation across variable such as age, sex and free school meal eligibility.

Local Authority data was aggregated to the regional and national level and checked against totals published for these geographic levels, Regional data was aggregated to the national level and checked against published totals.

### SPC internal comparisons

In total `r nrow(SPC_internal_comparison_summary)` combinations of SPC data sets and grouping variables were tested. 

Of these:

`r SPC_all_match_total` (`r SPC_all_match_perc`) matched across regions and nations.

`r SPC_regional_mismatch_count` (`r SPC_regional_mismatch_perc`) mismatched across regions.

`r SPC_national_mismatch_count` (`r SPC_national_mismatch_prec`) mismatched across nations.

`r SPC_regional_national_mismatch_count` (`r SPC_regional_national_mismatch_prec`) mismatched across regions and nations.


```{r mismatch-tables}

SPC_regional_mismatch_rows %>%
  select(Internal_comparison_result,Source_and_grouping) %>%
  bind_rows(data.frame(Internal_comparison_result = "", Source_and_grouping = "")) %>%
knitr::kable(col.names = c("Mismatch type",
                           "Mismatch location"),
             align = "lrrrr",
             caption = "Regional mismatches in SPC data")

SPC_national_mismatch_rows %>%
  select(Internal_comparison_result,Source_and_grouping) %>%
  bind_rows(data.frame(Internal_comparison_result = "", Source_and_grouping = "")) %>%
knitr::kable(col.names = c("Mismatch type",
                           "Mismatch location"),
             align = "lrrrr",
             caption = "National mismatches in SPC data")

SPC_regional_national_mismatch_rows %>%
  select(Internal_comparison_result,Source_and_grouping) %>%
  bind_rows(data.frame(Internal_comparison_result = "", Source_and_grouping = "")) %>%
knitr::kable(col.names = c("Mismatch type",
                           "Mismatch location"),
             align = "lrrrr",
             caption = "Regional and National mismatches in SPC data")

```


### SPA internal comparisons

Temporary text


